\section{Introduction}

The described approach is mainly inspired from experience gained by using
analogue sound synthesizers. While every analogue synthesizer has its own
unique sound based on the physical parts it is made of, most do share
their key concepts. Usually a limited number of oscillators generate
signals resembling---more or less---sine waves which are then modulated
by being combined with each other in different ways.

\texttt{SOUNDLAB}---an experimental implementation of the presented
approach---is designed to enable the user to explore ways of signal
combination. It does so by defining an embedded \textit{domain specific
language} which provides axioms that generate primitive signals and
axioms that combine arbitrary signals into new signals. The semantics of
the language are based on a signal interface agreed on by every
component. Furthermore \texttt{SOUNDLAB} allows the use of
\textit{Common Lisp's} means of abstraction to define compound signals
and signal combinators. Primitive as well as compound parts of the system
form a homogeneous group of objects defined by their shared interfaces,
which grant the system power and flexibility of a \textit{Lisp} system.

There are of course many free software implementations\footnote{See for
instance Overtone (\texttt{http://overtone.github.io}) and Csound
(\texttt{http://www.csounds.com}).} of signal synthesis systems with
programming language interfaces. \texttt{SOUNDLAB} is---when compared to
others---much simpler and entirely written and embedded in \textit{Common
Lisp}.

\texttt{SOUNDLAB} is free software licensed under the \textit{GNU AGPL}
and can be obtained at \texttt{http://mr.gy/software/soundlab/}.
